---
title: "436proj"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

This project details our analysis of the movie dataset that contains information from [Rotten Tomatos](https://www.rottentomatoes.com/) and [IMDB](http://www.imdb.com/) for a random sample of movies. The purpose of this project is to develop a multiple linear regression model to understand what attributes make a movie popular. In the meantime, learning something new about movies. 

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
library(corrplot)
```

### Load data

```{r load-data}
#load("movies.Rdata")

ss <- readr::read_csv('spotify_songs.csv')
```

## Part 1: Data

meta data here

## Part 2: Research question

is song popularity related to energy, key, etc?

## Part 3: Exploratory data analysis and feature selection

Abstracting the data of the above potential predictors for the model.

```{r}
movies_new <- movies %>% select(title, title_type, genre, runtime, imdb_rating, imdb_num_votes, critics_rating, critics_score, audience_rating, audience_score, best_pic_win, best_actor_win, best_actress_win, best_dir_win)


```

Look at the structure of the data

```{r}
#str(movies_new)

str(ss)
```

Summary statistics

```{r}
#summary(movies_new)

summary(ss)
```

I find there is one missing value, and decide to drop it. 

```{r}
#movies_new <- na.omit(movies_new)

ss <- na.omit(ss)
```

Part of this project is to use the model to predict a movie's audience score and this movie should not be part of the data. Therefore, I split the data into traning and testing, and there is only one row in the test set.

```{r}
#set.seed(2017)
#split <- sample(seq_len(nrow(movies_new)), size = floor(0.999 * nrow(movies_new)))
#train <- movies_new[split, ]
#test <- movies_new[-split, ]
#dim(train)

set.seed(2017)
split <- sample(seq_len(nrow(ss)), size = floor(0.999 * nrow(ss)))
train <- ss[split, ]
test <- ss[-split, ]
dim(train)
```

```{r}
dim(test)
```

### Histogram of Numeric variables

```{r}
#hist(train$audience_score)
#summary(train$audience_score)

hist(train$track_popularity)
summary(train$track_popularity)
```





p1 <- ggplot(aes(x=runtime), data=train) + 
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white', binwidth = 5) + ylab('percentage') + ggtitle('Run Time')
p2 <- ggplot(aes(x=imdb_rating), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white', binwidth = 0.2) + ylab('percentage') + ggtitle('IMDB rating')
p3 <- ggplot(aes(x=log10(imdb_num_votes)), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('log(IMDB number of votes)')
p4 <- ggplot(aes(x=critics_score), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white', binwidth = 2) + ylab('percentage') + ggtitle('Critics Score')
grid.arrange(p1, p2, p3, p4, ncol=2)

```{r message=FALSE, warning=FALSE}


p1 <- ggplot(aes(x=duration_ms), data=train) + 
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white', binwidth = 5) + ylab('percentage') + ggtitle('duration_ms')
p2 <- ggplot(aes(x=danceability), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white', binwidth = 0.2) + ylab('percentage') + ggtitle('danceability')
p3 <- ggplot(aes(x=valence), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('valence')
p4 <- ggplot(aes(x=tempo), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white', binwidth = 2) + ylab('percentage') + ggtitle('tempo')
grid.arrange(p1, p2, p3, p4, ncol=2)


```

Regression analysis: 

### Bar plot of categorical variables
p1 <- ggplot(aes(x=title_type), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('Title Type') + coord_flip()
p2 <- ggplot(aes(x=genre), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('Genre') + coord_flip()
p3 <- ggplot(aes(x=critics_rating), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('Critics Rating') + coord_flip()
p4 <- ggplot(aes(x=audience_rating), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('Audience Rating') + coord_flip()
grid.arrange(p1, p2, p3, p4, ncol=2)

```{r}

p1 <- ggplot(aes(x=key), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('Key') + coord_flip()
p2 <- ggplot(aes(x=playlist_genre), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('playlist_genre') + coord_flip()
p3 <- ggplot(aes(x=playlist_subgenre), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('playlist_subgenre') + coord_flip()
p4 <- ggplot(aes(x=track_artist), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('track artist') + coord_flip()
grid.arrange(p1, p2, p3, p4, ncol=2)
```



### Correlation between numerical variables
vars <- names(train) %in% c('runtime', 'imdb_rating', 'imdb_num_votes', 'critics_score')
selected_train <- train[vars]
corr.matrix <- cor(selected_train)
corrplot(corr.matrix, main="\n\nCorrelation Plot of numerical variables", method="number")

```{r}

vars <- names(train) %in% c('duration_ms', 'danceability', 'valence', 'tempo')
selected_train <- train[vars]
corr.matrix <- cor(selected_train)
corrplot(corr.matrix, main="\n\nCorrelation Plot of numerical variables", method="number")


```



### Correlation between categorical variables and audience score
boxplot(audience_score~critics_rating, data=train, main='Audience score vs. Critics rating', xlab='Critics Rating', ylab='Audience Score')
by(train$audience_score, train$critics_rating, summary)
boxplot(audience_score~audience_rating, data=train, main='Audience Score vs. Audience Rating', xlab='Audience rating', ylab='Audience Score')
by(train$audience_score, train$audience_rating, summary)
boxplot(audience_score~title_type, data=train, main='Audience score vs. Title type', xlab='Title_type', ylab='Audience Score')
by(train$audience_score, train$title_type, summary)
boxplot(audience_score~genre, data=train, main='Audience score vs. Genre', xlab='Genre', ylab='Audience score')
by(train$audience_score, train$genre, summary)

```{r}


boxplot(track_popularity~key, data=train, main='track_popularity vs. key', xlab='key', ylab='track_popularity')
by(train$track_popularity, train$key, summary)
boxplot(track_popularity~playlist_genre, data=train, main='track_popularity vs. genre', xlab='genre', ylab='track_popularity')
by(train$track_popularity, train$playlist_genre, summary)
boxplot(track_popularity~playlist_subgenre, data=train, main='track_popularity vs. subgenre', xlab='subgenre', ylab='track_popularity')
by(train$track_popularity, train$playlist_subgenre, summary)

```



## Part 4: Modeling

We will be using stepwise model forward selection method, we start with an empty model, then add variables one at a time until a parsimonious model is reached. 

From the following full model, we can see that imdb rating has the lowest p value and is the most correlated variable to our response variable. So we choose imdb rating as the first predictor.

full_model <- lm(audience_score~imdb_rating+title_type+genre+runtime+imdb_num_votes+critics_rating+audience_rating+best_pic_win+best_actor_win+best_actress_win+best_dir_win, data=train)
summary(full_model)

```{r}

#check order of predictors 
full_model <- lm(track_popularity~duration_ms+key+playlist_genre+danceability+valence+energy+playlist_subgenre, data=train)
summary(full_model)
```


```{r}
fit1 <- lm(track_popularity ~ danceability, data=train)
summary(fit1)
```
