---
title: "436model"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup

data from
model based on

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
library(corrplot)
```

### Load data

```{r load-data}
ss <- readr::read_csv('spotify_songs.csv')
```

## Part 1: Data

meta data here

## Part 2: Research question

is song popularity related to energy, key, etc?

## Part 3: Exploratory data analysis and feature selection

Abstracting the data of the above potential predictors for the model.

movies_new <- movies %>% select(title, title_type, genre, runtime, imdb_rating, imdb_num_votes, critics_rating, critics_score, audience_rating, audience_score, best_pic_win, best_actor_win, best_actress_win, best_dir_win)

```{r}

```

Look at the structure of the data

```{r}
str(ss)
```

Summary statistics

```{r}

summary(ss)
```

drop missing values

```{r}
ss <- na.omit(ss)
```



Split data into training and testing

```{r}
set.seed(2017)
split <- sample(seq_len(nrow(ss)), size = floor(0.999 * nrow(ss)))
train <- ss[split, ]
test <- ss[-split, ]
dim(train)
```

```{r}
dim(test)
```

### Histogram of Numeric variables

```{r}
hist(train$popularity)
summary(train$popularity)
```




```{r message=FALSE, warning=FALSE}
p1 <- ggplot(aes(x=duration_ms), data=train) + 
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('duration_ms')
p2 <- ggplot(aes(x=danceability), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('danceability')
p3 <- ggplot(aes(x=valence), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('valence')
p4 <- ggplot(aes(x=tempo), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('tempo')
p5 <- ggplot(aes(x=energy), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('energy')
p6 <- ggplot(aes(x=acousticness), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('acousticness')
p7 <- ggplot(aes(x=speechiness), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('speechiness')
p8 <- ggplot(aes(x=loudness), data=train) +
  geom_histogram(aes(y=100*(..count..)/sum(..count..)), color='black', fill='white') + ylab('percentage') + ggtitle('loudness')
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, ncol=2)
```
Bad spread of distribution in acousticness, do not include.
Bad spread of distribution in speechiness, do not include.

Regression analysis: 

### Bar plot of categorical variables


```{r}
p1 <- ggplot(aes(x=key), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('Key') + coord_flip()
p2 <- ggplot(aes(x=genre), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('genre') + coord_flip()
p4 <- ggplot(aes(x=artist), data=train) + geom_bar(aes(y=100*(..count..)/sum(..count..))) + ylab('percentage') +
  ggtitle('track artist') + coord_flip()
grid.arrange(p1, p2, p4, ncol=2)
```
Too many artists, do not include.


### Correlation between numerical variables


```{r}
vars <- names(train) %in% c('duration_ms', 'danceability', 'valence', 'tempo', 'energy', 'loudness', 'instrumentalness')
selected_train <- train[vars]
corr.matrix <- cor(selected_train)
corrplot(corr.matrix, main="\n\nCorrelation Plot of numerical variables", method="number")
```

High correlation between loudness and energy (0.65), so only include energy

### Correlation between categorical variables and audience score


```{r}
boxplot(popularity~key, data=train, main='popularity vs. key', xlab='key', ylab='popularity')
by(train$popularity, train$key, summary)
boxplot(popularity~genre, data=train, main='popularity vs. genre', xlab='genre', ylab='popularity')
by(train$popularity, train$genre, summary)
```
Average populairty varies by genre but not by key.
Therefore,
Genre does have correlation with audience score
Key does not have correlation with audience score


## Part 4: Modeling

```{r}
#check order of predictors 
full_model <- lm(popularity~duration_ms+tempo+genre+danceability+valence+energy+instrumentalness, data=train)
summary(full_model)
```


```{r}
fit1 <- lm(popularity ~ danceability, data=train)
summary(fit1)
```


```{r}
anova(full_model)
```

### Model diagnostics

```{r}
ggplot(data = full_model, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```

```{r}
ggplot(data = full_model, aes(x = .resid)) +
  geom_histogram(binwidth = 1, fill='white', color='black') +
  xlab("Residuals")
```

```{r}
ggplot(data = full_model, aes(sample = .resid)) +
  stat_qq()
```

## Part 5: Prediction

```{r}
newsong <- test %>% select(duration_ms, tempo, genre, danceability, valence, energy, instrumentalness)
predict(full_model, newsong)
```

```{r}
predict(full_model, newsong, interval = "prediction", level = 0.05)
```

```{r}
test$popularity
```
